---
title: "Chrotomyini brms holding tank"
author: "S.M. Smith"
date: "5/3/2022"
output: html_document
---

In WAIC diagnostics, each specimen has a "penalty term", p_WAIC, that lowers the total WAIC when it's super high. High p_WAIC indicates that the specimen is forcing the model to be overfit, and an overfit model is less good at 
```{r}
tibble(pareto_k = ch.26$criteria$loo$diagnostics$pareto_k,
       p_waic   = ch.26$criteria$waic$pointwise[, "p_waic"],
       specno      = pull(d, specno)) %>%
  ggplot(aes(x = pareto_k, y = p_waic)) +
  geom_vline(xintercept = .5, linetype = 2, color = "black", alpha = 1/2) +
  geom_point() +
  geom_text(data = . %>% filter(p_waic > 0.5),
            aes(x = pareto_k - 0.03, label = specno),
            hjust = 1) +
  labs(subtitle = "Gaussian model (b5.3)") +
  theme(legend.position = "none")
```

```{r}

```{r}
# BV.TV with only phylogenetic correlation structure:
ch.48 <- 
  brm(data = d, 
      family = gaussian,
      bvtv_s ~ 1 + (1|gr(phylo, cov = ch)),
      control = list(adapt_delta = 0.89), #inserted to decrease the number of divergent transitions here
      prior = c(
        prior(normal(0, 1), class = sd),
        prior(exponential(1), class = sigma)
        ),
      data2 = list(ch = ch),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      file = "G:\\My Drive\\Philippine rodents\\Chrotomyini_brms\\fits\\ch.48")
print(ch.48)
# NB! This model is primarily for comparing with other models - if you consider phylogenetic covariance structure only, does it allow you to predict bv.tv better than, or equally as well as, when you include mass/genus? I don't know that there's a good way to plot this. 
```


```{r}

p.36 <- 
  ggplot(d, aes(x=mass_s, y = bvtv_s, color = genus)) + 
  geom_point(pch = 16, cex = 4) +
  scale_color_manual(values = cols)
p.36

```

```

