---
title: "Mini phylo signal 06092022"
author: "S.M. Smith"
date: '2022-06-09'
output: 
  html_document: 
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
pacman::p_load(install = F, "ape", "bayesplot","BiocManager", "brms", "broom", "dagitty", "devtools", "flextable", "ggdark", "ggmcmc", "ggrepel", "ggtree", "gtools", "lattice","loo", "patchwork", "phytools","rcartocolor", "Rcpp", "remotes", "rstan", "StanHeaders", "statebins", "tidybayes", "viridis", "viridisLite", "pacman")
```

Load up Chrotomyini trabecular bone architecture (TBA) data and standardize variables:

```{r}
d <- read.csv(file = "G:\\My Drive\\Philippine rodents\\chrotomyini\\05062022 Philippine Murids segmentation parameters and morphological data - TBA data total BoneJ (full).csv", header = T)

d <- d[d$tribe=="chroto",c(1:2, 4:23)]

d <- 
  d %>% 
  mutate(bvtv = as.numeric(bvtv))

d <- 
  d %>%
  mutate(mass_s = rethinking::standardize(log10(mass_g)),
         elev_s = rethinking::standardize(elev), 
         bvtv_s = rethinking::standardize(bvtv),
         tbth_s = rethinking::standardize(tbth),
         tbsp_s = rethinking::standardize(tbsp),
         conn_s = rethinking::standardize(conn),
         cond_s = rethinking::standardize(m_connd),
         cond_s2 = rethinking::standardize(connd), 
         da_s = rethinking::standardize(da))

# remove C. gonzalesi and R. isarogensis, singletons:
d <- 
  d %>% 
  filter(taxon!="Chrotomys_gonzalesi") %>% 
  filter(taxon!="Rhynchomys_isarogensis")

# Make categorical vars into factors
d <- 
  d %>%
  mutate(loco = factor(loco), 
         hab_simp = factor(hab_simp),
         genus = factor(genus))

# Specify colors for plots:
cols = c("#86acca","#ab685b", "#3370a3", "#1f314c","#5a9fa8")
```

Load in phylogeny: 
REMEMBER: A <- ape::vcv.phylo(phylo), add corr = T if your tree is NOT SCALED TO 1. 

```{r}
ch.tre <- read.nexus(file = "G:\\My Drive\\Philippine rodents\\Chrotomys\\analysis\\SMS_PRUNED_and_COLLAPSED_03292022_OTUsrenamed_Rowsey_PhBgMCC_LzChrotomyini.nex")

ch <- ape::vcv.phylo(ch.tre, corr = T)

d <- 
  d %>% 
  mutate(phylo = taxon)

```

Estimate lambda: all four trabecular bone variables. All are by taxon (species-level) and include mass and phylo structure terms. Eg:

```{r}
ch.74.3 <- 
  brm(file = "G:\\My Drive\\Philippine rodents\\chrotomyini\\fits\\ch.74.3")
print(ch.74.3)
```

Calculating lambda:


```{r}
hyp <- "sd_phylo__Intercept^2 / (sd_phylo__Intercept^2 + sigma^2) = 0"

#BV.TV
h.bvtv <- hypothesis(ch.74.3, hyp, class = NULL)

#Tb.Th
ch.76 <- 
  brm(file = "G:\\My Drive\\Philippine rodents\\chrotomyini\\fits\\ch.76")
h.tbth <- hypothesis(ch.76, hyp, class = NULL)

#Tb.Sp
ch.77 <- 
  brm(file = "G:\\My Drive\\Philippine rodents\\chrotomyini\\fits\\ch.77")
h.tbsp <- hypothesis(ch.77, hyp, class = NULL)

#Conn.d
ch.78 <- 
  brm(file = "G:\\My Drive\\Philippine rodents\\chrotomyini\\fits\\ch.78")
h.cond <- hypothesis(ch.78, hyp, class = NULL)


bvtv.stpl <- ggplot() +
  geom_density(aes(x = h.bvtv$samples$H1), fill = "orange", alpha = 0.5) +
  theme_bw() +
  xlim(0,1) +
  labs(y = "density", x = "lambda: bone volume fraction")

tbth.stpl <- ggplot() +
  geom_density(aes(x = h.tbth$samples$H1), fill = "orange", alpha = 0.5) +
  theme_bw() +
  xlim(0,1) +
  labs(y = "density", x = "lambda: trabecular thickness")

tbsp.stpl <- ggplot() +
  geom_density(aes(x = h.tbsp$samples$H1), fill = "orange", alpha = 0.5) +
  theme_bw() +
  xlim(0,1) +
  labs(y = "density", x = "lambda: trabecular separation")

cond.stpl <- ggplot() +
  geom_density(aes(x = h.cond$samples$H1), fill = "orange", alpha = 0.5) +
  theme_bw() +
  xlim(0,1) +
  labs(y = "density", x = "lambda: connectivity density")
bvtv.stpl/tbsp.stpl|tbth.stpl/cond.stpl
```

All have peak density at low values for lambda except Tb.Th, which is pretty confused. Tb.Th is the most correlated with mass, so I checked on what the phylogenetic signal is in mass by itself:


```{r}
ch.80 <- 
  brm(file = "G:\\My Drive\\Philippine rodents\\chrotomyini\\fits\\ch.80")
h.mass <- hypothesis(ch.80, hyp, class = NULL)
mass.stpl <- ggplot() +
  geom_density(aes(x = h.mass$samples$H1), fill = "orange", alpha = 0.5) +
  theme_bw() +
  xlim(0,1) +
  labs(y = "density", x = "lambda: mass")
mass.stpl
```


Pretty high peak near 1. Is the phylo signal reflective of the effect that mass has on the metric? What if we calculate lambda on tbth without including mass? If that's the issue, I would expect that without mass as a predictor there is a very high phylo signal.

```{r}
ch.80.2 <-
  brm(data = d, 
      family = student,
      tbth_s ~ 0 + taxon + (1|gr(phylo, cov = ch)),
      control = list(adapt_delta = 0.98), #inserted to decrease the number of divergent transitions here
      prior = c(
        prior(gamma(2, 0.1), class = nu),
        prior(normal(0, 1), class = b),
        prior(normal(0, 1), class = sd),
        prior(exponential(1), class = sigma)
        ),
      data2 = list(ch = ch),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      file = "G:\\My Drive\\Philippine rodents\\chrotomyini\\fits\\ch.80.2")
print(ch.80.2)
h.thnomass <- hypothesis(ch.80.2, hyp, class = NULL)
thnomass.stpl <- ggplot() +
  geom_density(aes(x = h.thnomass$samples$H1), fill = "orange", alpha = 0.5) +
  theme_bw() +
  xlim(0,1) +
  labs(y = "density", x = "lambda: trabecular thickness, no-mass model")
thnomass.stpl
```

I think this result supports my idea about what's causing this. Check the other metrics:

```{r}
ch.80.3 <-
  brm(data = d, 
      family = student,
      bvtv_s ~ 0 + taxon + (1|gr(phylo, cov = ch)),
      control = list(adapt_delta = 0.98), #inserted to decrease the number of divergent transitions here
      prior = c(
        prior(gamma(2, 0.1), class = nu),
        prior(normal(0, 1), class = b),
        prior(normal(0, 1), class = sd),
        prior(exponential(1), class = sigma)
        ),
      data2 = list(ch = ch),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      file = "G:\\My Drive\\Philippine rodents\\chrotomyini\\fits\\ch.80.3")
h.bvnomass <- hypothesis(ch.80.3, hyp, class = NULL)
bvnomass.stpl <- ggplot() +
  geom_density(aes(x = h.bvnomass$samples$H1), fill = "orange", alpha = 0.5) +
  theme_bw() +
  xlim(0,1) +
  labs(y = "density", x = "lambda: bone volume fraction, no-mass model")
bvnomass.stpl
```

Ok so maybe that's the cause here. When you add mass to the Tb.Th model, the error structures are too similar to be able to tell if the error is from phylo or from mass, and there's very little other information there.  

Compare with conn.d, which has a negative correlation with mass, but the fit isn't great:

```{r}
ch.80.4 <-
  brm(data = d, 
      family = student,
      cond_s ~ 0 + taxon + (1|gr(phylo, cov = ch)),
      control = list(adapt_delta = 0.98), #inserted to decrease the number of divergent transitions here
      prior = c(
        prior(gamma(2, 0.1), class = nu),
        prior(normal(0, 1), class = b),
        prior(normal(0, 1), class = sd),
        prior(exponential(1), class = sigma)
        ),
      data2 = list(ch = ch),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      file = "G:\\My Drive\\Philippine rodents\\chrotomyini\\fits\\ch.80.4")
h.cdnomass <- hypothesis(ch.80.4, hyp, class = NULL)
cdnomass.stpl <- ggplot() +
  geom_density(aes(x = h.cdnomass$samples$H1), fill = "orange", alpha = 0.5) +
  theme_bw() +
  xlim(0,1) +
  labs(y = "density", x = "lambda: connectivity density, no-mass model")
cdnomass.stpl
```

So even though the slope of mass vs. conn.d is big, there is a lot more variance in the relationship. What about some pearson correlations?
```{r}
cor.test(d$mass_s, d$bvtv_s) # cor = 0.57, big CIs (0.39-0.72)

cor.test(d$mass_s, d$tbth_s) # cor = 0.93, v small CIs (0.89-0.95)

cor.test(d$mass_s, d$tbsp_s) # cor = 0.39, big CIs (0.16-0.57)

cor.test(d$mass_s, d$cond_s) # cor = -0.68, medium CIs (-0.79 - -0.52)

```


Can I do all the mass influence terms from the previous models?

```{r}
a <- mcmc_plot(ch.74.3, pars = "^b_mass")
b <- mcmc_plot(ch.76, pars = "^b_mass")
c <- mcmc_plot(ch.77, pars = "^b_mass")
d <- mcmc_plot(ch.78, pars = "^b_mass")

a <- a +
  xlim(-2,1.5)
b <- b +
  xlim(-2,1.5)
c <- c +
  xlim(-2,1.5)
d <- d +
  xlim(-2,1.5)

a/b/c/d

```

