---
title: "Comparing genus modeling"
author: "Stephanie M Smith"
date: '2022-04-20'
output: 
  html_document: 
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
pacman::p_load(install = F, "ape", "bayesplot","BiocManager", "brms", "broom", "dagitty", "devtools", "flextable", "ggdark", "ggmcmc", "ggrepel", "gtools", "lattice","loo", "patchwork", "rcartocolor", "Rcpp", "remotes", "rstan", "StanHeaders", "statebins", "tidybayes", "viridis", "viridisLite", "pacman")
```

Load up your guys and standardize:
```{r}
d <- read.csv(file = "G:\\My Drive\\Philippine rodents\\Chrotomyini_brms\\04082022 Philippine Murids segmentation parameters and morphological data - TBA data total BoneJ (full).csv", header = T)

d <- d[d$tribe=="chroto",c(1:2, 4:22)]

d <- 
  d %>% 
  mutate(bvtv = as.numeric(bvtv))

d <- 
  d %>%
  mutate(mass_s = rethinking::standardize(log10(mass_g)),
         elev_s = rethinking::standardize(elev), 
         bvtv_s = rethinking::standardize(bvtv),
         tbth_s = rethinking::standardize(tbth),
         tbsp_s = rethinking::standardize(tbsp),
         conn_s = rethinking::standardize(conn),
         cond_s = rethinking::standardize(connd), 
         da_s = rethinking::standardize(da))

# remove C. gonzalesi and R. isarogensis
d <- 
  d %>% 
  filter(taxon!="Chrotomys_gonzalesi") %>% 
  filter(taxon!="Rhynchomys_isarogensis")

d <- 
  d %>%
  mutate(loco = factor(loco), 
         hab_simp = factor(hab_simp),
         genus = factor(genus))
```

Load in phylogeny: 
REMEMBER: A <- ape::vcv.phylo(phylo), add corr = T if your tree is NOT SCALED TO 1. 
```{r}
ch.tre <- read.nexus(file = "G:\\My Drive\\Philippine rodents\\Chrotomys\\analysis\\SMS_PRUNED_and_COLLAPSED_03292022_OTUsrenamed_Rowsey_PhBgMCC_LzChrotomyini.nex")

ch <- ape::vcv.phylo(ch.tre, corr = T)

d <- 
  d %>% 
  mutate(phylo = taxon)

```

In the "genusasloco" file, I used genus (loc proxy) as a global/fixed effect:
~genus + (1|species) : population level effect 

What if I do it as a group level effect/random effect?
~(1|genus) + (1|species)

Let's try that and see how it affects the results. 

```{r}
# Load up a previous model for comparison: Genus only with no phylogeny or anything.
ch.24 <- 
  brm(data = d, 
      family = gaussian,
      bvtv_s ~ 0 + genus,
      prior = c(prior(normal(0, 1), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 5,
      file = "G:\\My Drive\\Philippine rodents\\Chrotomyini_brms\\fits\\ch.24")
print(ch.24)

# Comparison model: genus only but as a group-level effect. 
ch.30 <- 
  brm(data = d, 
      family = gaussian,
      bvtv_s ~ 0 + (1|genus), 
      prior = c(#prior(normal(0, 1), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 5,
      file = "G:\\My Drive\\Philippine rodents\\Chrotomyini_brms\\fits\\ch.30")
print(ch.30)
plot(ch.30)
mcmc_plot(ch.30)

```

So this doesn't give you estimates for every genus. To be honest I'm not sure what exactly this is telling me other than the amount of error attributable to genus. I should try to find out exactly what this notation means. 
