---
title: "Chrotomyini genus-separated models"
author: "S.M. Smith"
date: "4/8/2022"
output: 
  html_document: 
    keep_md: true
---
A handy reminder: a new %>% is made with ctrl+shift+m.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
pacman::p_load(install = F, "ape", "bayesplot","BiocManager", "brms", "broom", "dagitty", "devtools", "flextable", "GGally", "ggdag", "ggdark", "ggmcmc", "ggrepel", "ggthemes",  "ghibli", "gtools", "lattice","loo", "patchwork", "psych", "rcartocolor", "Rcpp", "remotes", "rstan", "StanHeaders", "statebins", "tidybayes", "viridis", "viridisLite", "wesanderson", "pacman")
```

Load up your guys and standardize:
```{r}
d <- read.csv(file = "G:\\My Drive\\Philippine rodents\\Chrotomyini_brms\\04082022 Philippine Murids segmentation parameters and morphological data - TBA data total BoneJ (full).csv", header = T)

d <- d[d$tribe=="chroto",c(1:2, 4:22)]

d <- 
  d %>% 
  mutate(bvtv = as.numeric(bvtv))

d <- 
  d %>%
  mutate(mass_s = rethinking::standardize(log10(mass_g)),
         elev_s = rethinking::standardize(elev), 
         bvtv_s = rethinking::standardize(bvtv),
         tbth_s = rethinking::standardize(tbth),
         tbsp_s = rethinking::standardize(tbsp),
         conn_s = rethinking::standardize(conn),
         cond_s = rethinking::standardize(connd), 
         da_s = rethinking::standardize(da))

# remove C. gonzalesi and R. isarogensis
d <- 
  d %>% 
  filter(taxon!="Chrotomys_gonzalesi") %>% 
  filter(taxon!="Rhynchomys_isarogensis")

cols = c("#86acca","#ab685b", "#3370a3", "#1f314c","#5a9fa8")

d <- 
  d %>%
  mutate(loco = factor(loco), 
         hab_simp = factor(hab_simp),
         genus = factor(genus))
pairs(d[,c(22,24,25,26,28)])

dcov <- cov(d[,c(22,24,25,26,29)]) 
dcor <- cov2cor(dcov)
levelplot(dcor, col.regions = mako(100))
```


Load in phylogeny: 
REMEMBER: A <- ape::vcv.phylo(phylo), add corr = T if your tree is NOT SCALED TO 1. 
```{r}
ch.tre <- read.nexus(file = "G:\\My Drive\\Philippine rodents\\Chrotomys\\analysis\\SMS_PRUNED_and_COLLAPSED_03292022_OTUsrenamed_Rowsey_PhBgMCC_LzChrotomyini.nex")
plot(ch.tre)

ch <- ape::vcv.phylo(ch.tre, corr = T)

d <- 
  d %>% 
  mutate(phylo = taxon)

```

Plot of Chrotomys only:
```{r}
chbvf<- d %>%
  filter(genus == "Chrotomys") %>% 
  ggplot()+
  geom_boxplot(aes(x=taxon, y = bvtv), lwd = 1.5, fill = "#3370A3", color = "#1F314C")

chtbth<- d %>%
  filter(genus == "Chrotomys") %>% 
  ggplot()+
  geom_boxplot(aes(x=taxon, y = tbth), lwd = 1.5, fill = "#3370A3", color = "#1F314C")

chtbsp<- d %>%
  filter(genus == "Chrotomys") %>% 
  ggplot()+
  geom_boxplot(aes(x=taxon, y = tbsp), lwd = 1.5, fill = "#3370A3", color = "#1F314C")

chbvf|chtbth

```

Model of Chrotomys only: 
```{r}
# Chrotomys only data set
chronly <- d %>%
  filter(genus == "Chrotomys")

# Chrotomys only tree
chron_tre <-keep.tip(ch.tre,ch.tre$tip.label[grep("Chrotomys",ch.tre$tip.label)])
chcov <- ape::vcv.phylo(chron_tre, corr = T)

# by taxon only
ch.45 <- 
  brm(data = chronly, 
      family = gaussian,
      bvtv_s ~ 0 + taxon,
      prior = c(prior(normal(0, 1), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      file = "G:\\My Drive\\Philippine rodents\\Chrotomyini_brms\\fits\\ch.45")
print(ch.45)
mcmc_plot(ch.45)

# by taxon plus mass
ch.46 <- 
  brm(data = chronly, 
      family = gaussian,
      bvtv_s ~ 0 + taxon + mass_s,
      prior = c(prior(normal(0, 1), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      file = "G:\\My Drive\\Philippine rodents\\Chrotomyini_brms\\fits\\ch.46")
print(ch.46)
mcmc_plot(ch.46)

# by taxon plus mass plus phy/intrasp
ch.47 <-
  brm(data = chronly, 
      family = gaussian,
      bvtv_s ~ 0 + taxon + mass_s + (1|gr(phylo, cov = chcov)) + (1|taxon),
      control = list(adapt_delta = 0.97), #inserted to decrease the number of divergent transitions here
      prior = c(
        prior(normal(0, 1), class = b, coef = taxonChrotomys_whiteheadi),
        prior(normal(0, 1), class = b, coef = taxonChrotomys_silaceus),
        prior(normal(0, 1), class = b, coef = taxonChrotomys_mindorensis),
        prior(normal(0, 1), class = b, coef = mass_s),
        prior(normal(0, 1), class = sd),
        prior(exponential(1), class = sigma)
        ),
      data2 = list(chcov = chcov),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      file = "G:\\My Drive\\Philippine rodents\\Chrotomyini_brms\\fits\\ch.47")
print(ch.47)
mcmc_plot(ch.47)

hyp <- "sd_phylo__Intercept^2 / (sd_phylo__Intercept^2 + sd_taxon__Intercept^2 + sigma^2) = 0"
chrot.bvtv <- hypothesis(ch.47, hyp, class = NULL)
chrot.bvtv.pl <- ggplot() +
  stat_halfeye(aes(x = chrot.bvtv$samples$H1), fill = "red", alpha = 0.5) +
  theme_bw() +
  xlim(0,1) +
  labs(y = "density", x = "lambda: bone volume fraction")
chrot.bvtv.pl
```

```{r}
ch.47 %>%
  gather_draws(b_taxonChrotomys_whiteheadi,b_taxonChrotomys_silaceus, b_taxonChrotomys_mindorensis) %>%
  compare_levels(.value, by = .variable) %>%
  ungroup() %>%
  #mutate(loco = reorder(.variable, .value)) %>%
  ggplot(aes(y = .variable, x = .value)) +
  stat_halfeye(.width = .89) +
  geom_vline(xintercept = 0, linetype = "dashed")
```


ch.24: compare bvtv across genera, no phylo or nothin
```{r}
ch.24 <- 
  brm(data = d, 
      family = gaussian,
      bvtv_s ~ 0 + genus,
      prior = c(prior(normal(0, 1), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 5,
      file = "G:\\My Drive\\Philippine rodents\\Chrotomyini_brms\\fits\\ch.24")
print(ch.24)
mcmc_plot(ch.24)
ch.24 %>%
  gather_draws(b_genusApomys,b_genusArchboldomys, b_genusChrotomys, b_genusRhynchomys, b_genusSoricomys) %>%
  ggplot(aes(y = .variable, x = .value)) +
  stat_halfeye(aes(fill = .variable), 
               point_fill = "#000000", 
               shape = 21, 
               point_size = 3, 
               point_color = "#FFFFFF",
               interval_size = 7,
               interval_color = "grey40",
               .width = .89) +
  scale_fill_manual(values = cols)+
  theme(legend.position = "none")+
  geom_vline(xintercept = 0, linetype = "dashed")+
  ggtitle(label = "Bone volume fraction by genus, no size or phylo")
```
ch.25: compare bvtv across genera, mass but no phylo

```{r}
ch.25 <- 
  brm(data = d, 
      family = gaussian,
      bvtv_s ~ 0 + genus + mass_s,
      prior = c(prior(normal(0, 1), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 5,
      file = "G:\\My Drive\\Philippine rodents\\Chrotomyini_brms\\fits\\ch.25")
print(ch.25)
mcmc_plot(ch.25, "^b_", regex = T)
ch.25 %>%
  gather_draws(b_genusApomys,b_genusArchboldomys, b_genusChrotomys, b_genusRhynchomys, b_genusSoricomys) %>%
  ggplot(aes(y = .variable, x = .value)) +
  stat_halfeye(aes(fill = .variable), 
               point_fill = "#000000", 
               shape = 21, 
               point_size = 3, 
               point_color = "#FFFFFF",
               interval_size = 7,
               interval_color = "grey40",
               .width = .89) +
  scale_fill_manual(values = cols)+
  theme(legend.position = "none")+
  geom_vline(xintercept = 0, linetype = "dashed")+
  ggtitle(label = "Bone volume fraction by genus, no phylo")
```
...this is not at all what I expected?


ch.26: bvtv by genus + mass, Include phylo and intraspec:
```{r}
ch.26 <-
  brm(data = d, 
      family = gaussian,
      bvtv_s ~ 0 + genus + mass_s + (1|gr(phylo, cov = ch)) + (1|taxon),
      control = list(adapt_delta = 0.85), #inserted to decrease the number of divergent transitions here
      prior = c(
        prior(normal(0, 1), class = b, coef = genusApomys),
        prior(normal(0, 1), class = b, coef = genusArchboldomys),
        prior(normal(0, 1), class = b, coef = genusChrotomys),
        prior(normal(0, 1), class = b, coef = genusRhynchomys),
        prior(normal(0, 1), class = b, coef = genusSoricomys),
        prior(normal(0, 1), class = b, coef = mass_s),
        prior(normal(0, 1), class = sd),
        prior(exponential(1), class = sigma)
        ),
      data2 = list(ch = ch),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      file = "G:\\My Drive\\Philippine rodents\\Chrotomyini_brms\\fits\\ch.26")

ch.26 <- add_criterion(ch.26, "waic")
ch.26 <- add_criterion(ch.26, "loo")

print(ch.26)
mcmc_plot(ch.26, "^b_", regex = T)


```

Halfeye plot: 
```{r}

ch.26 %>%
  gather_draws(b_genusApomys,b_genusArchboldomys, b_genusChrotomys, b_genusRhynchomys, b_genusSoricomys) %>%
  ggplot(aes(y = .variable, x = .value)) +
  stat_halfeye(aes(fill = .variable), 
               point_fill = "#000000", 
               shape = 21, 
               point_size = 3, 
               point_color = "#FFFFFF",
               interval_size = 7,
               interval_color = "grey40",
               .width = .89) +
  scale_fill_manual(values = cols)+
  theme(legend.position = "none")+
  geom_vline(xintercept = 0, linetype = "dashed")+
  ggtitle(label = "Bone volume fraction by genus")

ch.26 %>%
  gather_draws(b_genusApomys,b_genusArchboldomys, b_genusChrotomys, b_genusRhynchomys, b_genusSoricomys) %>%
  compare_levels(.value, by = .variable) %>%
  ungroup() %>%
  mutate(loco = reorder(.variable, .value)) %>%
  ggplot(aes(y = .variable, x = .value)) +
  stat_halfeye() +
  geom_vline(xintercept = 0, linetype = "dashed")
```

Combined halfeye plot of gradual addition of features to the model:
```{r}
nomass_halfeye <- ch.24 %>%
  gather_draws(b_genusApomys,b_genusArchboldomys, b_genusChrotomys, b_genusRhynchomys, b_genusSoricomys) %>%
  ggplot(aes(y = .variable, x = .value)) +
  stat_halfeye(aes(fill = .variable), 
               point_fill = "#000000", 
               shape = 21, 
               point_size = 7, 
               point_color = "#FFFFFF",
               interval_size = 15,
               interval_color = "grey40",
               .width = .89) +
  scale_fill_manual(values = cols)+
  geom_vline(xintercept = 0, linetype = "dashed")+
  theme(legend.position = "none")+
  ggtitle(label = "BVTV by genus only")

nophy_halfeye <- ch.25 %>%
  gather_draws(b_genusApomys,b_genusArchboldomys, b_genusChrotomys, b_genusRhynchomys, b_genusSoricomys) %>%
  ggplot(aes(y = .variable, x = .value)) +
  stat_halfeye(aes(fill = .variable), 
               point_fill = "#000000", 
               shape = 21, 
               point_size = 7, 
               point_color = "#FFFFFF",
               interval_size = 15,
               interval_color = "grey40",
               .width = .89) +
  scale_fill_manual(values = cols)+
  geom_vline(xintercept = 0, linetype = "dashed")+
  ggtitle(label = "BVTV, mass only")+
  theme(legend.position = "none",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank())

all_halfeye <- ch.26 %>%
  gather_draws(b_genusApomys,b_genusArchboldomys, b_genusChrotomys, b_genusRhynchomys, b_genusSoricomys) %>%
  ggplot(aes(y = .variable, x = .value)) +
  stat_halfeye(aes(fill = .variable), 
               point_fill = "#000000", 
               shape = 21, 
               point_size = 7, 
               point_color = "#FFFFFF",
               interval_size = 15,
               interval_color = "grey40",
               .width = .89) +
  scale_fill_manual(values = cols)+
  geom_vline(xintercept = 0, linetype = "dashed")+
  ggtitle(label = "BVTV, mass/phylo")+
  theme(legend.position = "none",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank())

nomass_halfeye | nophy_halfeye | all_halfeye
```

```{r}
loo(ch.26) # only one k above 0.5
tibble(k   = ch.26$criteria$loo$diagnostics$pareto_k,
       row = 1:67,
       specimen = paste(d$specno[row], d$taxon[row])) %>% 
  arrange(desc(k))
```

ch.27: tbth by genus, Include phylo and intraspec:
```{r}
ch.27 <-
  brm(data = d, 
      family = gaussian,
      tbth_s ~ 0 + genus + mass_s + (1|gr(phylo, cov = ch)) + (1|taxon),
      control = list(adapt_delta = 0.95), #inserted to decrease the number of divergent transitions here
      prior = c(
        prior(normal(0, 1), class = b, coef = genusApomys),
        prior(normal(0, 1), class = b, coef = genusArchboldomys),
        prior(normal(0, 1), class = b, coef = genusChrotomys),
        prior(normal(0, 1), class = b, coef = genusRhynchomys),
        prior(normal(0, 1), class = b, coef = genusSoricomys),
        prior(normal(0, 1), class = b, coef = mass_s),
        prior(normal(0, 1), class = sd),
        prior(exponential(1), class = sigma)
        ),
      data2 = list(ch = ch),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      file = "G:\\My Drive\\Philippine rodents\\Chrotomyini_brms\\fits\\ch.27")

print(ch.27)
mcmc_plot(ch.27)
ch.27 <- add_criterion(ch.27, "waic")
ch.27 <- add_criterion(ch.27, "loo")

loo(ch.27) #only one k above 0.5

ch.27 %>%
  gather_draws(b_genusApomys,b_genusArchboldomys, b_genusChrotomys, b_genusRhynchomys, b_genusSoricomys) %>%
  ggplot(aes(y = .variable, x = .value)) +
  stat_halfeye(aes(fill = .variable), 
               point_fill = "#000000", 
               shape = 21, 
               point_size = 7, 
               point_color = "#FFFFFF",
               interval_size = 15,
               interval_color = "grey40",
               .width = .89) +
  scale_fill_manual(values = cols)+
  geom_vline(xintercept = 0, linetype = "dashed")+
  #ggtitle(label = "tbth by genus")+
  theme(legend.position = "none",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank())
```

ch.28: da by genus, Include phylo and intraspec:
```{r}
ch.28 <-
  brm(data = d, 
      family = gaussian,
      da_s ~ 0 + genus + mass_s + (1|gr(phylo, cov = ch)) + (1|taxon),
      control = list(adapt_delta = 0.95), #inserted to decrease the number of divergent transitions here
      prior = c(
        prior(normal(0, 1), class = b, coef = genusApomys),
        prior(normal(0, 1), class = b, coef = genusArchboldomys),
        prior(normal(0, 1), class = b, coef = genusChrotomys),
        prior(normal(0, 1), class = b, coef = genusRhynchomys),
        prior(normal(0, 1), class = b, coef = genusSoricomys),
        prior(normal(0, 1), class = b, coef = mass_s),
        prior(normal(0, 1), class = sd),
        prior(exponential(1), class = sigma)
        ),
      data2 = list(ch = ch),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      file = "G:\\My Drive\\Philippine rodents\\Chrotomyini_brms\\fits\\ch.28")

print(ch.28)
mcmc_plot(ch.28)

ch.28 %>%
  gather_draws(b_genusApomys,b_genusArchboldomys, b_genusChrotomys, b_genusRhynchomys, b_genusSoricomys) %>%
  ggplot(aes(y = .variable, x = .value)) +
  stat_halfeye(aes(fill = .variable), 
               point_fill = "#000000", 
               shape = 21, 
               point_size = 7, 
               point_color = "#FFFFFF",
               interval_size = 15,
               interval_color = "grey40",
               .width = .89) +
  scale_fill_manual(values = cols)+
  geom_vline(xintercept = 0, linetype = "dashed")+
  #ggtitle(label = "da by genus")+
  theme(legend.position = "none",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank())

```

A very striking difference in Soricomys!!! 

ch.29: tbsp by genus, Include phylo and intraspec:
```{r}
ch.29 <-
  brm(data = d, 
      family = gaussian,
      tbsp_s ~ 0 + genus + mass_s + (1|gr(phylo, cov = ch)) + (1|taxon),
      control = list(adapt_delta = 0.95), #inserted to decrease the number of divergent transitions here
      prior = c(
        prior(normal(0, 1), class = b, coef = genusApomys),
        prior(normal(0, 1), class = b, coef = genusArchboldomys),
        prior(normal(0, 1), class = b, coef = genusChrotomys),
        prior(normal(0, 1), class = b, coef = genusRhynchomys),
        prior(normal(0, 1), class = b, coef = genusSoricomys),
        prior(normal(0, 1), class = b, coef = mass_s),
        prior(normal(0, 1), class = sd),
        prior(exponential(1), class = sigma)
        ),
      data2 = list(ch = ch),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      file = "G:\\My Drive\\Philippine rodents\\Chrotomyini_brms\\fits\\ch.29")

print(ch.29)
mcmc_plot(ch.29)

ch.29 %>%
  gather_draws(b_genusApomys,b_genusArchboldomys, b_genusChrotomys, b_genusRhynchomys, b_genusSoricomys) %>%
  ggplot(aes(y = .variable, x = .value)) +
  stat_halfeye(aes(fill = .variable), 
               point_fill = "#000000", 
               shape = 21, 
               point_size = 7, 
               point_color = "#FFFFFF",
               interval_size = 15,
               interval_color = "grey40",
               .width = .89) +
  scale_fill_manual(values = cols)+
  geom_vline(xintercept = 0, linetype = "dashed")+
  #ggtitle(label = "tbsp by genus")+
  theme(legend.position = "none",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank())

```


Compile effects of mass across all different metrics:
```{r}
gm.bvtv <- ch.26 %>% 
  gather_draws(b_mass_s) %>% 
  mutate(metric = "bvtv")

gm.tbth <- ch.27 %>% 
  gather_draws(b_mass_s) %>% 
  mutate(metric = "tbth")

gm.tbsp <- ch.29 %>% 
  gather_draws(b_mass_s) %>% 
  mutate(metric = "tbsp")

gm.da <- ch.28 %>% 
  gather_draws(b_mass_s) %>% 
  mutate(metric = "da")

gmass_b_all <- bind_rows(gm.bvtv, gm.tbth, gm.tbsp, gm.da)

gmass.pl <-   gmass_b_all %>% 
  ggplot(aes(y = metric, x = .value)) +
  stat_halfeye(aes(fill = .variable), 
               point_fill = "#000000", 
               shape = 21, 
               point_size = 3, 
               point_color = "#FFFFFF",
               interval_size = 7,
               interval_color = "grey40",
               .width = .89) +
  scale_fill_manual(values = "#CB2F7F")+
  theme(legend.position = "none")+
  geom_vline(xintercept = 0, linetype = "dashed")+
  ggtitle(label = "estimated effect of mass")
gmass.pl
```

Compile phylogenetic sd across all metrics:
```{r}
gph.bvtv <- ch.26 %>% 
  gather_draws(sd_phylo__Intercept) %>% 
  mutate(metric = "bvtv")

gph.tbth <- ch.27 %>% 
  gather_draws(sd_phylo__Intercept) %>% 
  mutate(metric = "tbth")

gph.tbsp <- ch.29 %>% 
  gather_draws(sd_phylo__Intercept) %>% 
  mutate(metric = "tbsp")

gph.da <- ch.28 %>% 
  gather_draws(sd_phylo__Intercept) %>% 
  mutate(metric = "da")

gphylo_sd_all <- bind_rows(gph.bvtv, gph.tbth, gph.tbsp, gph.da)

gph.pl <-   gphylo_sd_all %>% 
  ggplot(aes(y = metric, x = .value)) +
  stat_halfeye(aes(fill = .variable), 
               point_fill = "#000000", 
               shape = 21, 
               point_size = 3, 
               point_color = "#FFFFFF",
               interval_size = 7,
               interval_color = "grey40",
               .width = .89) +
  scale_fill_manual(values = "#CB2F7F")+
  theme(legend.position = "none")+
  geom_vline(xintercept = 0, linetype = "dashed")+
  ggtitle(label = "estimated phylo sd (NOT lambda!)")
gph.pl
```

Do one for phylogenetic signal:
```{r}
hyp <- "sd_phylo__Intercept^2 / (sd_phylo__Intercept^2 + sd_taxon__Intercept^2 + sigma^2) = 0"
h.bvtv <- hypothesis(ch.26, hyp, class = NULL)
h.tbth <- hypothesis(ch.27, hyp, class = NULL)
h.tbsp <- hypothesis(ch.29, hyp, class = NULL)
h.da <- hypothesis(ch.28, hyp, class = NULL)

ph.bvtv.pl <- ggplot() +
  stat_halfeye(aes(x = h.bvtv$samples$H1), fill = "red", alpha = 0.5) +
  theme_bw() +
  xlim(0,1) +
  labs(y = "density", x = "lambda: bone volume fraction")

ph.tbth.pl <- ggplot() +
  stat_halfeye(aes(x = h.tbth$samples$H1), fill = "orange", alpha = 0.5)+
  theme_bw() +
  xlim(0,1)+
  labs(y = "density", x = "lambda: trabecular thickness")

ph.tbsp.pl <- ggplot() +
  stat_halfeye(aes(x = h.tbsp$samples$H1), fill = "turquoise", alpha = 0.5)+
  theme_bw() +
  xlim(0,1) +
  labs(y = "density", x = "lambda: trabecular spacing")

ph.da.pl <- ggplot()+
  stat_halfeye(aes(x = h.da$samples$H1), fill = "grey40", alpha = 0.5) +
  theme_bw() +
  xlim(0,1)+
  labs(y = "density", x = "lambda: degree of anisotropy")

ph.bvtv.pl/ph.tbth.pl/ph.tbsp.pl/ph.da.pl
```

----------------------------------
Let's do some analyses that don't include mass as a covariate, but DO include phylogeny. 

Bone volume fraction:

```{r}
ch.44 <- 
  brm(data = d, 
      family = gaussian,
      bvtv_s ~ 0 + genus + (1|gr(phylo, cov = ch)) + (1|taxon),
      control = list(adapt_delta = 0.89), #inserted to decrease the number of divergent transitions here
      prior = c(
        prior(normal(0, 1), class = b, coef = genusApomys),
        prior(normal(0, 1), class = b, coef = genusArchboldomys),
        prior(normal(0, 1), class = b, coef = genusChrotomys),
        prior(normal(0, 1), class = b, coef = genusRhynchomys),
        prior(normal(0, 1), class = b, coef = genusSoricomys),
        prior(normal(0, 1), class = sd),
        prior(exponential(1), class = sigma)
        ),
      data2 = list(ch = ch),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      file = "G:\\My Drive\\Philippine rodents\\Chrotomyini_brms\\fits\\ch.44")

ch.44 <- add_criterion(ch.44, "waic")
ch.44 <- add_criterion(ch.44, "loo")

print(ch.44)
mcmc_plot(ch.44, "^b_", regex = T)
```

Compare: phy/mass (ch.26, all_halfeye), phy only (ch.44), mass only (ch.25,nophy_halfeye).

```{r}
phyonly_halfeye <- ch.44 %>%
  gather_draws(b_genusApomys,b_genusArchboldomys, b_genusChrotomys, b_genusRhynchomys, b_genusSoricomys) %>%
  ggplot(aes(y = .variable, x = .value)) +
  stat_halfeye(aes(fill = .variable), 
               point_fill = "#000000", 
               shape = 21, 
               point_size = 7, 
               point_color = "#FFFFFF",
               interval_size = 15,
               interval_color = "grey40",
               .width = .89) +
  scale_fill_manual(values = cols)+
  geom_vline(xintercept = 0, linetype = "dashed")+
  ggtitle(label = "BVTV, phy/int only")+
  theme(legend.position = "none",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank())
nomass_halfeye|phyonly_halfeye|nophy_halfeye|all_halfeye
```

It's probably worth doing the other metrics later but for the moment I want to do some intrageneric Chrotomys stuff.

