---
title: "Chrotomys analysis compilation - genera, with Student priors"
author: "S.M. Smith"
date: "5/2/2022"
output: 
  html_document: 
    keep_md: true
---

NEW ANALYSES using Student dist priors. My distributions need to be less easily surprised because TBA is pretty variable. 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
pacman::p_load(install = F, "ape", "bayesplot","BiocManager", "brms", "broom", "dagitty", "devtools", "flextable", "ggdark", "ggmcmc", "ggrepel", "gtools", "lattice","loo", "patchwork", "rcartocolor", "Rcpp", "remotes", "rstan", "StanHeaders", "statebins", "tidybayes", "viridis", "viridisLite", "pacman")
```

Load up Chrotomyini trabecular bone architecture (TBA) data and standardize variables:
```{r}
d <- read.csv(file = "G:\\My Drive\\Philippine rodents\\chrotomyini\\05062022 Philippine Murids segmentation parameters and morphological data - TBA data total BoneJ (full).csv", header = T)

d <- d[d$tribe=="chroto",c(1:2, 4:23)]

d <- 
  d %>% 
  mutate(bvtv = as.numeric(bvtv))

d <- 
  d %>%
  mutate(mass_s = rethinking::standardize(log10(mass_g)),
         elev_s = rethinking::standardize(elev), 
         bvtv_s = rethinking::standardize(bvtv),
         tbth_s = rethinking::standardize(tbth),
         tbsp_s = rethinking::standardize(tbsp),
         conn_s = rethinking::standardize(conn),
         cond_s = rethinking::standardize(m_connd),
         cond_s2 = rethinking::standardize(connd), 
         da_s = rethinking::standardize(da))

# remove C. gonzalesi and R. isarogensis, singletons:
d <- 
  d %>% 
  filter(taxon!="Chrotomys_gonzalesi") %>% 
  filter(taxon!="Rhynchomys_isarogensis")

# Make categorical vars into factors
d <- 
  d %>%
  mutate(loco = factor(loco), 
         hab_simp = factor(hab_simp),
         genus = factor(genus))

# Specify colors for plots:
cols = c("#86acca","#ab685b", "#3370a3", "#1f314c","#5a9fa8")
```

Load in phylogeny: 
REMEMBER: A <- ape::vcv.phylo(phylo), add corr = T if your tree is NOT SCALED TO 1. 
```{r}
ch.tre <- read.nexus(file = "G:\\My Drive\\Philippine rodents\\Chrotomys\\analysis\\SMS_PRUNED_and_COLLAPSED_03292022_OTUsrenamed_Rowsey_PhBgMCC_LzChrotomyini.nex")

ch <- ape::vcv.phylo(ch.tre, corr = T)

d <- 
  d %>% 
  mutate(phylo = taxon)
```

#######################################################################
#### Comparisons of several models for each trabecular bone metric ####
#######################################################################

Same analyses as before, but this time with Student's dist priors. For reference:
https://solomonkurz.netlify.app/post/2019-02-02-robust-linear-regression-with-student-s-t-distribution/
https://solomonkurz.netlify.app/post/2019-02-10-bayesian-robust-correlations-with-brms-and-why-you-should-love-student-s-t/

######################################
#### Bone Volume Fraction (BV.TV) ####
######################################


```{r}
# BV.TV by mass
ch.70 <- 
  brm(data = d, 
      family = student,
      bvtv_s ~ 1+mass_s,
      prior = c(prior(normal(0, 1), class = b),
                prior(gamma(2, 0.1), class = nu),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 5,
      file = "G:\\My Drive\\Philippine rodents\\chrotomyini\\fits\\ch.70")
print(ch.70)
```

A scatterplot of thTe above model, with genera color-coded (the same color code to be used for the remainder of this document):
```{r}
range(d$mass_s)
nd <- tibble(mass_s = seq(from = -2, to = 1.75, length.out = 67))

fitted(ch.70,
       newdata = nd) %>% 
  data.frame() %>% 
  bind_cols(nd) %>% 
  
  # plot
  ggplot(aes(x = mass_s)) +
  geom_smooth(aes(y = Estimate, ymin = Q2.5, ymax = Q97.5),
              stat = "identity",
              alpha = 1/5, size = 1, color = "black") +
  geom_point(data = d, aes(y = bvtv_s, color = genus), size = 4)+
  scale_color_manual(values = cols)+
  xlim(min(nd), max(nd))+
  labs(x = "Mass in grams (standardized)",
       y = "bone volume fraction (standardized)") 
```

Testing across metrics:

```{r}
# By genus only
ch.71 <- 
  brm(data = d, 
      family = student,
      bvtv_s ~ 0 + genus,
      prior = c(prior(normal(0, 1), class = b),
                prior(gamma(2, 0.1), class = nu),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      file = "G:\\My Drive\\Philippine rodents\\chrotomyini\\fits\\ch.71")
print(ch.71)
```


```{r}
# By genus and mass
ch.72 <- 
  brm(data = d, 
      family = student,
      bvtv_s ~ 0 + genus + mass_s,
      prior = c(prior(normal(0, 1), class = b),
                prior(gamma(2, 0.1), class = nu),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      file = "G:\\My Drive\\Philippine rodents\\chrotomyini\\fits\\ch.72")
print(ch.72)
```


```{r}
# By genus, mass, and phylogeny

ch.74.1 <-
  brm(data = d, 
      family = student,
      bvtv_s ~ 0 + genus + mass_s + (1|gr(phylo, cov = ch)),
      control = list(adapt_delta = 0.85), #inserted to decrease the number of divergent transitions here
      prior = c(
        prior(gamma(2, 0.1), class = nu),
        prior(normal(0, 1), class = b, coef = genusApomys),
        prior(normal(0, 1), class = b, coef = genusArchboldomys),
        prior(normal(0, 1), class = b, coef = genusChrotomys),
        prior(normal(0, 1), class = b, coef = genusRhynchomys),
        prior(normal(0, 1), class = b, coef = genusSoricomys),
        prior(normal(0, 1), class = b, coef = mass_s),
        prior(normal(0, 1), class = sd),
        prior(exponential(1), class = sigma)
        ),
      data2 = list(ch = ch),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      file = "G:\\My Drive\\Philippine rodents\\chrotomyini\\fits\\ch.74.1")
print(ch.74.1)

#By species, for funsies
ch.74.3 <-
  brm(data = d, 
      family = student,
      bvtv_s ~ 0 + taxon + mass_s + (1|gr(phylo, cov = ch)),
      control = list(adapt_delta = 0.85), #inserted to decrease the number of divergent transitions here
      prior = c(
        prior(gamma(2, 0.1), class = nu),
        prior(normal(0, 1), class = b),
        prior(normal(0, 1), class = sd),
        prior(exponential(1), class = sigma)
        ),
      data2 = list(ch = ch),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      file = "G:\\My Drive\\Philippine rodents\\chrotomyini\\fits\\ch.74.3")
print(ch.74.3)

```

A four-part plot including all of the models:

```{r}
ch.71_halfeye <- ch.71 %>%
  gather_draws(b_genusApomys,b_genusArchboldomys, b_genusChrotomys, b_genusRhynchomys, b_genusSoricomys) %>%
  mutate(.variable = str_replace_all(.variable, "b_genus", "")) %>%
  ggplot(aes(y = .variable, x = .value)) +
  stat_halfeye(aes(fill = .variable), 
               point_fill = "#000000", 
               shape = 21, 
               point_size = 3, 
               point_color = "#FFFFFF",
               interval_size = 10,
               interval_color = "grey40",
               .width = .89) +
  scale_fill_manual(values = cols)+
  geom_vline(xintercept = 0, linetype = "dashed")+
  theme(legend.position = "none", 
        plot.title = element_text(size = 9))+
  labs(x = "BV.TV", y = "genus")+
  ggtitle(label = "BV.TV by genus only")

ch.72_halfeye <- ch.72 %>%
  gather_draws(b_genusApomys,b_genusArchboldomys, b_genusChrotomys, b_genusRhynchomys, b_genusSoricomys) %>%
  mutate(.variable = str_replace_all(.variable, "b_genus", "")) %>%
  ggplot(aes(y = .variable, x = .value)) +
  stat_halfeye(aes(fill = .variable), 
               point_fill = "#000000", 
               shape = 21, 
               point_size = 3, 
               point_color = "#FFFFFF",
               interval_size = 10,
               interval_color = "grey40",
               .width = .89) +
  scale_fill_manual(values = cols)+
  geom_vline(xintercept = 0, linetype = "dashed")+
  ggtitle(label = "BV.TV by genus/mass")+
  labs(x = "BV.TV", y = "genus")+
  theme(legend.position = "none", 
        plot.title = element_text(size = 9))

ch.74.1_halfeye <- ch.74.1 %>%
  gather_draws(b_genusApomys,b_genusArchboldomys, b_genusChrotomys, b_genusRhynchomys, b_genusSoricomys) %>%
  mutate(.variable = str_replace_all(.variable, "b_genus", "")) %>%
  ggplot(aes(y = .variable, x = .value)) +
  stat_halfeye(aes(fill = .variable), 
               point_fill = "#000000", 
               shape = 21, 
               point_size = 3, 
               point_color = "#FFFFFF",
               interval_size = 10,
               interval_color = "grey40",
               .width = .89) +
  scale_fill_manual(values = cols)+
  geom_vline(xintercept = 0, linetype = "dashed")+
  ggtitle(label = "BV.TV by genus/mass/phylo")+
  labs(x = "BV.TV", y = "genus")+
  theme(legend.position = "none", 
        plot.title = element_text(size = 9))

ch.71_halfeye|ch.72_halfeye|ch.74.1_halfeye

```

Here are the difference distributions between pairs of genera for the model including only genus, mass, and phylo as predictors:
```{r}
ch.74.1 %>%
  gather_draws(b_genusApomys,b_genusArchboldomys, b_genusChrotomys, b_genusRhynchomys, b_genusSoricomys) %>%
  compare_levels(.value, by = .variable) %>%
  mutate(.variable = str_replace_all(.variable, "b_genus", "")) %>%
  ungroup() %>%
  mutate(loco = reorder(.variable, .value)) %>%
  ggplot(aes(y = .variable, x = .value)) +
  stat_halfeye() +
  geom_vline(xintercept = 0, linetype = "dashed") +
  ggtitle(label = "Between-genus BV.TV difference distributions")+
  labs(x = "difference", y = "comparison")
```

```{r}
hyp <- "sd_phylo__Intercept^2 / (sd_phylo__Intercept^2 + sigma^2) = 0"
h.bvtv <- hypothesis(ch.74.1, hyp, class = NULL)
h.bvtv
```

Lambda = 0.56
```{r}
ph.bvtv.pl <- ggplot() +
  geom_density(aes(x = h.bvtv$samples$H1), fill = "red", alpha = 0.5) +
  theme_bw() +
  xlim(0,1) +
  labs(y = "density", x = "lambda: bone volume fraction")
ph.bvtv.pl
```


```{r}
ch.71 <-  add_criterion(ch.71, c("waic", "loo")) # mass only
ch.72 <-  add_criterion(ch.72, c("waic", "loo")) # genus only 
ch.74.1 <-  add_criterion(ch.74.1, c("waic", "loo")) # genus mass phylogeny

loo(ch.71) # genus only : 0 Pareto K > 0.5
loo(ch.72) # genus and mass : 0 Pareto K > 0.5
loo(ch.74.1) # genus mass phylogeny : 0 Pareto K > 0.5
```


######################################
#### Trabecular Thickness (Tb.Th) ####
######################################

A model without genus, mass as the only predictor:

```{r}
# Tb.Th by mass
ch.81 <- 
  brm(data = d, 
      family = student,
      tbth_s ~ 1+mass_s,
      prior = c(prior(normal(0, 1), class = b),
                prior(gamma(2, 0.1), class = nu),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      file = "G:\\My Drive\\Philippine rodents\\chrotomyini\\fits\\ch.81")
print(ch.81)
```

A scatterplot of the above model, with genera color-coded:
```{r}
ch.81.pl <- fitted(ch.81,
       newdata = nd) %>% 
  data.frame() %>% 
  bind_cols(nd) %>% 
  
  # plot
  ggplot(aes(x = mass_s)) +
  geom_smooth(aes(y = Estimate, ymin = Q2.5, ymax = Q97.5),
              stat = "identity",
              alpha = 1/5, size = 1, color = "black") +
  geom_point(data = d, aes(y = tbth_s, color = genus), size = 4)+
  scale_color_manual(values = cols)+
  labs(x = "Mass in grams (standardized)",
       y = "trabecular thickness (standardized)") 
ch.81.pl
```

```{r}
# By genus only
ch.82 <- 
  brm(data = d, 
      family = student,
      tbth_s ~ 0 + genus,
      prior = c(prior(normal(0, 1), class = b),
                prior(gamma(2, 0.1), class = nu),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      file = "G:\\My Drive\\Philippine rodents\\chrotomyini\\fits\\ch.82")
print(ch.82)
```

```{r}
# By genus and mass
ch.83 <- 
  brm(data = d, 
      family = student,
      tbth_s ~ 0 + genus + mass_s,
      prior = c(prior(normal(0, 1), class = b),
                prior(gamma(2, 0.1), class = nu),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      file = "G:\\My Drive\\Philippine rodents\\chrotomyini\\fits\\ch.83")
print(ch.83)
```


```{r}
# By genus, mass, and phylogeny

ch.84 <-
  brm(data = d, 
      family = student,
      tbth_s ~ 0 + genus + mass_s + (1|gr(phylo, cov = ch)),
      control = list(adapt_delta = 0.85), #inserted to decrease the number of divergent transitions here
      prior = c(
        prior(gamma(2, 0.1), class = nu),
        prior(normal(0, 1), class = b, coef = genusApomys),
        prior(normal(0, 1), class = b, coef = genusArchboldomys),
        prior(normal(0, 1), class = b, coef = genusChrotomys),
        prior(normal(0, 1), class = b, coef = genusRhynchomys),
        prior(normal(0, 1), class = b, coef = genusSoricomys),
        prior(normal(0, 1), class = b, coef = mass_s),
        prior(normal(0, 1), class = sd),
        prior(exponential(1), class = sigma)
        ),
      data2 = list(ch = ch),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      file = "G:\\My Drive\\Philippine rodents\\chrotomyini\\fits\\ch.84")
print(ch.84)
```

A four-part plot including all of the models:

```{r}
ch.82_halfeye <- ch.82 %>%
  gather_draws(b_genusApomys,b_genusArchboldomys, b_genusChrotomys, b_genusRhynchomys, b_genusSoricomys) %>%
  mutate(.variable = str_replace_all(.variable, "b_genus", "")) %>%
  ggplot(aes(y = .variable, x = .value)) +
  stat_halfeye(aes(fill = .variable), 
               point_fill = "#000000", 
               shape = 21, 
               point_size = 3, 
               point_color = "#FFFFFF",
               interval_size = 10,
               interval_color = "grey40",
               .width = .89) +
  scale_fill_manual(values = cols)+
  geom_vline(xintercept = 0, linetype = "dashed")+
  theme(legend.position = "none", 
        plot.title = element_text(size = 9))+
  labs(x = "Tb.Th", y = "genus")+
  ggtitle(label = "Tb.Th by genus only")

ch.83_halfeye <- ch.83 %>%
  gather_draws(b_genusApomys,b_genusArchboldomys, b_genusChrotomys, b_genusRhynchomys, b_genusSoricomys) %>%
  mutate(.variable = str_replace_all(.variable, "b_genus", "")) %>%
  ggplot(aes(y = .variable, x = .value)) +
  stat_halfeye(aes(fill = .variable), 
               point_fill = "#000000", 
               shape = 21, 
               point_size = 3, 
               point_color = "#FFFFFF",
               interval_size = 10,
               interval_color = "grey40",
               .width = .89) +
  scale_fill_manual(values = cols)+
  geom_vline(xintercept = 0, linetype = "dashed")+
  ggtitle(label = "Tb.Th by genus/mass")+
  labs(x = "Tb.Th", y = "genus")+
  theme(legend.position = "none", 
        plot.title = element_text(size = 9))

ch.84_halfeye <- ch.84 %>%
  gather_draws(b_genusApomys,b_genusArchboldomys, b_genusChrotomys, b_genusRhynchomys, b_genusSoricomys) %>%
  mutate(.variable = str_replace_all(.variable, "b_genus", "")) %>%
  ggplot(aes(y = .variable, x = .value)) +
  stat_halfeye(aes(fill = .variable), 
               point_fill = "#000000", 
               shape = 21, 
               point_size = 3, 
               point_color = "#FFFFFF",
               interval_size = 10,
               interval_color = "grey40",
               .width = .89) +
  scale_fill_manual(values = cols)+
  geom_vline(xintercept = 0, linetype = "dashed")+
  ggtitle(label = "Tb.Th by genus/mass/phylo")+
  labs(x = "Tb.Th", y = "genus")+
  theme(legend.position = "none", 
        plot.title = element_text(size = 9))

ch.82_halfeye|ch.83_halfeye|ch.84_halfeye

```

```{r}
hyp <- "sd_phylo__Intercept^2 / (sd_phylo__Intercept^2 + sigma^2) = 0"
h.tbth <- hypothesis(ch.84, hyp, class = NULL)
h.tbth
```

Lambda = 0.7
```{r}
ph.bvtv.pl <- ggplot() +
  geom_density(aes(x = h.tbth$samples$H1), fill = "red", alpha = 0.5) +
  theme_bw() +
  xlim(0,1) +
  labs(y = "density", x = "lambda: trabecular thickness")
ph.bvtv.pl
```


#####################################
#### Trabecular Separation Tb.Sp ####
#####################################

A model without genus, mass as the only predictor:

```{r}
# Tb.Th by mass
ch.85<- 
  brm(data = d, 
      family = student,
      tbsp_s ~ 1+mass_s,
      prior = c(prior(normal(0, 1), class = b),
                prior(gamma(2, 0.1), class = nu),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      file = "G:\\My Drive\\Philippine rodents\\chrotomyini\\fits\\ch.85")
print(ch.85)
```

A scatterplot of the above model, with genera color-coded:
```{r}
ch.85.pl <- fitted(ch.85,
       newdata = nd) %>% 
  data.frame() %>% 
  bind_cols(nd) %>% 
  
  # plot
  ggplot(aes(x = mass_s)) +
  geom_smooth(aes(y = Estimate, ymin = Q2.5, ymax = Q97.5),
              stat = "identity",
              alpha = 1/5, size = 1, color = "black") +
  geom_point(data = d, aes(y = tbsp_s, color = genus), size = 4)+
  scale_color_manual(values = cols)+
  labs(x = "Mass in grams (standardized)",
       y = "trabecular separation (standardized)") 
ch.85.pl
```

```{r}
# By genus only
ch.86 <- 
  brm(data = d, 
      family = student,
      tbsp_s ~ 0 + genus,
      prior = c(prior(normal(0, 1), class = b),
                prior(gamma(2, 0.1), class = nu),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      file = "G:\\My Drive\\Philippine rodents\\chrotomyini\\fits\\ch.86")
print(ch.86)
```

```{r}
# By genus and mass
ch.87 <- 
  brm(data = d, 
      family = student,
      tbsp_s ~ 0 + genus + mass_s,
      prior = c(prior(normal(0, 1), class = b),
                prior(gamma(2, 0.1), class = nu),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      file = "G:\\My Drive\\Philippine rodents\\chrotomyini\\fits\\ch.87")
print(ch.87)
```


```{r}
# By genus, mass, and phylogeny

ch.88 <-
  brm(data = d, 
      family = student,
      tbsp_s ~ 0 + genus + mass_s + (1|gr(phylo, cov = ch)),
      control = list(adapt_delta = 0.85), #inserted to decrease the number of divergent transitions here
      prior = c(
        prior(gamma(2, 0.1), class = nu),
        prior(normal(0, 1), class = b, coef = genusApomys),
        prior(normal(0, 1), class = b, coef = genusArchboldomys),
        prior(normal(0, 1), class = b, coef = genusChrotomys),
        prior(normal(0, 1), class = b, coef = genusRhynchomys),
        prior(normal(0, 1), class = b, coef = genusSoricomys),
        prior(normal(0, 1), class = b, coef = mass_s),
        prior(normal(0, 1), class = sd),
        prior(exponential(1), class = sigma)
        ),
      data2 = list(ch = ch),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      file = "G:\\My Drive\\Philippine rodents\\chrotomyini\\fits\\ch.88")
print(ch.88)
```

A four-part plot including all of the models:

```{r}
ch.86_halfeye <- ch.86 %>%
  gather_draws(b_genusApomys,b_genusArchboldomys, b_genusChrotomys, b_genusRhynchomys, b_genusSoricomys) %>%
  mutate(.variable = str_replace_all(.variable, "b_genus", "")) %>%
  ggplot(aes(y = .variable, x = .value)) +
  stat_halfeye(aes(fill = .variable), 
               point_fill = "#000000", 
               shape = 21, 
               point_size = 3, 
               point_color = "#FFFFFF",
               interval_size = 10,
               interval_color = "grey40",
               .width = .89) +
  scale_fill_manual(values = cols)+
  geom_vline(xintercept = 0, linetype = "dashed")+
  theme(legend.position = "none", 
        plot.title = element_text(size = 9))+
  labs(x = "Tb.Sp", y = "genus")+
  ggtitle(label = "Tb.Sp by genus only")

ch.87_halfeye <- ch.87 %>%
  gather_draws(b_genusApomys,b_genusArchboldomys, b_genusChrotomys, b_genusRhynchomys, b_genusSoricomys) %>%
  mutate(.variable = str_replace_all(.variable, "b_genus", "")) %>%
  ggplot(aes(y = .variable, x = .value)) +
  stat_halfeye(aes(fill = .variable), 
               point_fill = "#000000", 
               shape = 21, 
               point_size = 3, 
               point_color = "#FFFFFF",
               interval_size = 10,
               interval_color = "grey40",
               .width = .89) +
  scale_fill_manual(values = cols)+
  geom_vline(xintercept = 0, linetype = "dashed")+
  ggtitle(label = "Tb.Sp by genus/mass")+
  labs(x = "Tb.Sp", y = "genus")+
  theme(legend.position = "none", 
        plot.title = element_text(size = 9))

ch.88_halfeye <- ch.88 %>%
  gather_draws(b_genusApomys,b_genusArchboldomys, b_genusChrotomys, b_genusRhynchomys, b_genusSoricomys) %>%
  mutate(.variable = str_replace_all(.variable, "b_genus", "")) %>%
  ggplot(aes(y = .variable, x = .value)) +
  stat_halfeye(aes(fill = .variable), 
               point_fill = "#000000", 
               shape = 21, 
               point_size = 3, 
               point_color = "#FFFFFF",
               interval_size = 10,
               interval_color = "grey40",
               .width = .89) +
  scale_fill_manual(values = cols)+
  geom_vline(xintercept = 0, linetype = "dashed")+
  ggtitle(label = "Tb.Sp by genus/mass/phylo")+
  labs(x = "Tb.Sp", y = "genus")+
  theme(legend.position = "none", 
        plot.title = element_text(size = 9))

ch.86_halfeye|ch.87_halfeye|ch.88_halfeye

```

```{r}
hyp <- "sd_phylo__Intercept^2 / (sd_phylo__Intercept^2 + sigma^2) = 0"
h.tbsp <- hypothesis(ch.88, hyp, class = NULL)
h.tbsp
```

Lambda = 0.27
```{r}
ph.tbsp.pl <- ggplot() +
  geom_density(aes(x = h.tbsp$samples$H1), fill = "red", alpha = 0.5) +
  theme_bw() +
  xlim(0,1) +
  labs(y = "density", x = "lambda: trabecular separation")
ph.tbsp.pl
```


#####################################
#### Connectivity density Conn.D ####
#####################################

A model without genus, mass as the only predictor:

```{r}
# Conn.D by mass
ch.89<- 
  brm(data = d, 
      family = student,
      cond_s ~ 1+mass_s,
      prior = c(prior(normal(0, 1), class = b),
                prior(gamma(2, 0.1), class = nu),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      file = "G:\\My Drive\\Philippine rodents\\chrotomyini\\fits\\ch.89")
print(ch.89)
```

A scatterplot of the above model, with genera color-coded:
```{r}
ch.89.pl <- fitted(ch.89,
       newdata = nd) %>% 
  data.frame() %>% 
  bind_cols(nd) %>% 
  
  # plot
  ggplot(aes(x = mass_s)) +
  geom_smooth(aes(y = Estimate, ymin = Q2.5, ymax = Q97.5),
              stat = "identity",
              alpha = 1/5, size = 1, color = "black") +
  geom_point(data = d, aes(y = cond_s, color = genus), size = 4)+
  scale_color_manual(values = cols)+
  labs(x = "Mass in grams (standardized)",
       y = "connectivity density (standardized)") 
ch.89.pl
```

```{r}
# By genus only
ch.90 <- 
  brm(data = d, 
      family = student,
      cond_s ~ 0 + genus,
      prior = c(prior(normal(0, 1), class = b),
                prior(gamma(2, 0.1), class = nu),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      file = "G:\\My Drive\\Philippine rodents\\chrotomyini\\fits\\ch.90")
print(ch.90)
```

```{r}
# By genus and mass
ch.91 <- 
  brm(data = d, 
      family = student,
      cond_s ~ 0 + genus + mass_s,
      prior = c(prior(normal(0, 1), class = b),
                prior(gamma(2, 0.1), class = nu),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      file = "G:\\My Drive\\Philippine rodents\\chrotomyini\\fits\\ch.91")
print(ch.91)
```


```{r}
# By genus, mass, and phylogeny

ch.92 <-
  brm(data = d, 
      family = student,
      cond_s ~ 0 + genus + mass_s + (1|gr(phylo, cov = ch)),
      control = list(adapt_delta = 0.85), #inserted to decrease the number of divergent transitions here
      prior = c(
        prior(gamma(2, 0.1), class = nu),
        prior(normal(0, 1), class = b, coef = genusApomys),
        prior(normal(0, 1), class = b, coef = genusArchboldomys),
        prior(normal(0, 1), class = b, coef = genusChrotomys),
        prior(normal(0, 1), class = b, coef = genusRhynchomys),
        prior(normal(0, 1), class = b, coef = genusSoricomys),
        prior(normal(0, 1), class = b, coef = mass_s),
        prior(normal(0, 1), class = sd),
        prior(exponential(1), class = sigma)
        ),
      data2 = list(ch = ch),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      file = "G:\\My Drive\\Philippine rodents\\chrotomyini\\fits\\ch.92")
print(ch.92)
```

A four-part plot including all of the models:

```{r}
ch.90_halfeye <- ch.90 %>%
  gather_draws(b_genusApomys,b_genusArchboldomys, b_genusChrotomys, b_genusRhynchomys, b_genusSoricomys) %>%
  mutate(.variable = str_replace_all(.variable, "b_genus", "")) %>%
  ggplot(aes(y = .variable, x = .value)) +
  stat_halfeye(aes(fill = .variable), 
               point_fill = "#000000", 
               shape = 21, 
               point_size = 3, 
               point_color = "#FFFFFF",
               interval_size = 10,
               interval_color = "grey40",
               .width = .89) +
  scale_fill_manual(values = cols)+
  geom_vline(xintercept = 0, linetype = "dashed")+
  theme(legend.position = "none", 
        plot.title = element_text(size = 9))+
  labs(x = "Conn.D", y = "genus")+
  ggtitle(label = "Conn.D by genus only")

ch.91_halfeye <- ch.91 %>%
  gather_draws(b_genusApomys,b_genusArchboldomys, b_genusChrotomys, b_genusRhynchomys, b_genusSoricomys) %>%
  mutate(.variable = str_replace_all(.variable, "b_genus", "")) %>%
  ggplot(aes(y = .variable, x = .value)) +
  stat_halfeye(aes(fill = .variable), 
               point_fill = "#000000", 
               shape = 21, 
               point_size = 3, 
               point_color = "#FFFFFF",
               interval_size = 10,
               interval_color = "grey40",
               .width = .89) +
  scale_fill_manual(values = cols)+
  geom_vline(xintercept = 0, linetype = "dashed")+
  ggtitle(label = "Conn.D by genus/mass")+
  labs(x = "Conn.D", y = "genus")+
  theme(legend.position = "none", 
        plot.title = element_text(size = 9))

ch.92_halfeye <- ch.92 %>%
  gather_draws(b_genusApomys,b_genusArchboldomys, b_genusChrotomys, b_genusRhynchomys, b_genusSoricomys) %>%
  mutate(.variable = str_replace_all(.variable, "b_genus", "")) %>%
  ggplot(aes(y = .variable, x = .value)) +
  stat_halfeye(aes(fill = .variable), 
               point_fill = "#000000", 
               shape = 21, 
               point_size = 3, 
               point_color = "#FFFFFF",
               interval_size = 10,
               interval_color = "grey40",
               .width = .89) +
  scale_fill_manual(values = cols)+
  geom_vline(xintercept = 0, linetype = "dashed")+
  ggtitle(label = "Conn.D by genus/mass/phylo")+
  labs(x = "Conn.D", y = "genus")+
  theme(legend.position = "none", 
        plot.title = element_text(size = 9))

ch.90_halfeye|ch.91_halfeye|ch.92_halfeye

```

```{r}
hyp <- "sd_phylo__Intercept^2 / (sd_phylo__Intercept^2 + sigma^2) = 0"
h.cond <- hypothesis(ch.92, hyp, class = NULL)
h.cond
```

Lambda = 0.31
```{r}
ph.cond.pl <- ggplot() +
  geom_density(aes(x = h.cond$samples$H1), fill = "red", alpha = 0.5) +
  theme_bw() +
  xlim(0,1) +
  labs(y = "density", x = "lambda: connectivity density")
ph.cond.pl
```

#### Phylogenetic signal in mass by itself:
```{r}
ch.93<- 
  brm(data = d, 
      family = student,
      mass_s ~0 +genus+(1|gr(phylo, cov = ch)),
      prior = c(
        prior(gamma(2, 0.1), class = nu),
        prior(normal(0, 1), class = b, coef = genusApomys),
        prior(normal(0, 1), class = b, coef = genusArchboldomys),
        prior(normal(0, 1), class = b, coef = genusChrotomys),
        prior(normal(0, 1), class = b, coef = genusRhynchomys),
        prior(normal(0, 1), class = b, coef = genusSoricomys),
        prior(normal(0, 1), class = sd),
        prior(exponential(1), class = sigma)
        ),
      data2 = list(ch = ch),
      file = "G:\\My Drive\\Philippine rodents\\chrotomyini\\fits\\ch.93")
print(ch.93)
```

```{r}
h.mass <- hypothesis(ch.93, hyp, class = NULL)
h.mass
```

Lambda = 0.83.

```{r}
ph.mass.pl <- ggplot() +
  geom_density(aes(x = h.mass$samples$H1), fill = "red", alpha = 0.5) +
  theme_bw() +
  xlim(0,1) +
  labs(y = "density", x = "lambda: mass")
ph.mass.pl
```

```{r}
ch.94<- 
  brm(data = d, 
      family = student,
      bvtv_s ~0 +genus+(1|gr(phylo, cov = ch)),
      prior = c(
        prior(gamma(2, 0.1), class = nu),
        prior(normal(0, 1), class = b, coef = genusApomys),
        prior(normal(0, 1), class = b, coef = genusArchboldomys),
        prior(normal(0, 1), class = b, coef = genusChrotomys),
        prior(normal(0, 1), class = b, coef = genusRhynchomys),
        prior(normal(0, 1), class = b, coef = genusSoricomys),
        prior(normal(0, 1), class = sd),
        prior(exponential(1), class = sigma)
        ),
      data2 = list(ch = ch),
      file = "G:\\My Drive\\Philippine rodents\\chrotomyini\\fits\\ch.94")
print(ch.94)
h.94 <- hypothesis(ch.94, hyp, class = NULL)
h.94
ph.94.pl <- ggplot() +
  geom_density(aes(x = h.94$samples$H1), fill = "red", alpha = 0.5) +
  theme_bw() +
  xlim(0,1) +
  labs(y = "density", x = "lambda: bvtv")
ph.94.pl
```

```{r}
ch.95<- 
  brm(data = d, 
      family = student,
      tbth_s ~0 +genus+(1|gr(phylo, cov = ch)),
      prior = c(
        prior(gamma(2, 0.1), class = nu),
        prior(normal(0, 1), class = b, coef = genusApomys),
        prior(normal(0, 1), class = b, coef = genusArchboldomys),
        prior(normal(0, 1), class = b, coef = genusChrotomys),
        prior(normal(0, 1), class = b, coef = genusRhynchomys),
        prior(normal(0, 1), class = b, coef = genusSoricomys),
        prior(normal(0, 1), class = sd),
        prior(exponential(1), class = sigma)
        ),
      data2 = list(ch = ch),
      file = "G:\\My Drive\\Philippine rodents\\chrotomyini\\fits\\ch.95")
print(ch.95)
h.95 <- hypothesis(ch.95, hyp, class = NULL)
h.95
ph.95.pl <- ggplot() +
  geom_density(aes(x = h.95$samples$H1), fill = "red", alpha = 0.5) +
  theme_bw() +
  xlim(0,1) +
  labs(y = "density", x = "lambda: tbth")
ph.95.pl
```

```{r}
ch.96<- 
  brm(data = d, 
      family = student,
      tbsp_s ~0 +genus+(1|gr(phylo, cov = ch)),
      prior = c(
        prior(gamma(2, 0.1), class = nu),
        prior(normal(0, 1), class = b, coef = genusApomys),
        prior(normal(0, 1), class = b, coef = genusArchboldomys),
        prior(normal(0, 1), class = b, coef = genusChrotomys),
        prior(normal(0, 1), class = b, coef = genusRhynchomys),
        prior(normal(0, 1), class = b, coef = genusSoricomys),
        prior(normal(0, 1), class = sd),
        prior(exponential(1), class = sigma)
        ),
      data2 = list(ch = ch),
      file = "G:\\My Drive\\Philippine rodents\\chrotomyini\\fits\\ch.96")
print(ch.96)
h.96 <- hypothesis(ch.96, hyp, class = NULL)
h.96
ph.96.pl <- ggplot() +
  geom_density(aes(x = h.96$samples$H1), fill = "red", alpha = 0.5) +
  theme_bw() +
  xlim(0,1) +
  labs(y = "density", x = "lambda: tbsp")
ph.96.pl
```

```{r}
ch.97<- 
  brm(data = d, 
      family = student,
      cond_s ~0 +genus+(1|gr(phylo, cov = ch)),
      prior = c(
        prior(gamma(2, 0.1), class = nu),
        prior(normal(0, 1), class = b, coef = genusApomys),
        prior(normal(0, 1), class = b, coef = genusArchboldomys),
        prior(normal(0, 1), class = b, coef = genusChrotomys),
        prior(normal(0, 1), class = b, coef = genusRhynchomys),
        prior(normal(0, 1), class = b, coef = genusSoricomys),
        prior(normal(0, 1), class = sd),
        prior(exponential(1), class = sigma)
        ),
      data2 = list(ch = ch),
      file = "G:\\My Drive\\Philippine rodents\\chrotomyini\\fits\\ch.97")
print(ch.97)
h.97 <- hypothesis(ch.97, hyp, class = NULL)
h.97
ph.97.pl <- ggplot() +
  geom_density(aes(x = h.97$samples$H1), fill = "red", alpha = 0.5) +
  theme_bw() +
  xlim(0,1) +
  labs(y = "density", x = "lambda: conn.d")
ph.97.pl
```

