---
title: "Habitat effects"
author: "S.M. Smith"
date: "4/21/2022"
output: 
  html_document: 
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
pacman::p_load(install = F, "ape", "bayesplot","BiocManager", "brms", "broom", "dagitty", "devtools", "flextable", "ggdark", "ggmcmc", "ggrepel", "gtools", "lattice","loo", "patchwork", "rcartocolor", "Rcpp", "remotes", "rstan", "StanHeaders", "statebins", "tidybayes", "viridis", "viridisLite", "pacman")
```

Load up your guys and standardize:
```{r}
d <- read.csv(file = "G:\\My Drive\\Philippine rodents\\Chrotomyini_brms\\04082022 Philippine Murids segmentation parameters and morphological data - TBA data total BoneJ (full).csv", header = T)

d <- d[d$tribe=="chroto",c(1:2, 4:22)]

d <- 
  d %>% 
  mutate(bvtv = as.numeric(bvtv))

d <- 
  d %>%
  mutate(mass_s = rethinking::standardize(log10(mass_g)),
         elev_s = rethinking::standardize(elev), 
         bvtv_s = rethinking::standardize(bvtv),
         tbth_s = rethinking::standardize(tbth),
         tbsp_s = rethinking::standardize(tbsp),
         conn_s = rethinking::standardize(conn),
         cond_s = rethinking::standardize(connd), 
         da_s = rethinking::standardize(da))

# remove C. gonzalesi and R. isarogensis
d <- 
  d %>% 
  filter(taxon!="Chrotomys_gonzalesi") %>% 
  filter(taxon!="Rhynchomys_isarogensis")

d <- 
  d %>%
  mutate(loco = factor(loco), 
         hab_simp = factor(hab_simp),
         genus = factor(genus))
```

Load in phylogeny: 
REMEMBER: A <- ape::vcv.phylo(phylo), add corr = T if your tree is NOT SCALED TO 1. 
```{r}
ch.tre <- read.nexus(file = "G:\\My Drive\\Philippine rodents\\Chrotomys\\analysis\\SMS_PRUNED_and_COLLAPSED_03292022_OTUsrenamed_Rowsey_PhBgMCC_LzChrotomyini.nex")

ch <- ape::vcv.phylo(ch.tre, corr = T)

d <- 
  d %>% 
  mutate(phylo = taxon)

```

Alternate version of the data set excluding low-number habs:
```{r}
dd <- d 

dd <- 
  dd %>% 
  filter(hab_simp != "low_mont", 
         hab_simp != "mont_mos") %>% 
  droplevels(dd$hab_simp)
```

A brief enumeration of how many specimens fall into each habitat:
```{r}
d[which(d$hab_simp=="montane"),2] # 16 specimens
d[which(d$hab_simp=="mossy"),2] # 37 specimens
d[which(d$hab_simp=="lowland"),] # 7 specimens

# colors for plots later:
habcol <- c("#a2614d", "#2d5e43", "#8daf8e")
habmorecol <- c("#ed3945","#a2614d","#d99436","#2d5e43", "#8daf8e")
```

A base model looking at the relationship between mass and bv.tv:
```{r}
ch.36 <- 
  brm(data = d, 
      family = gaussian,
      bvtv_s ~ 1+mass_s,
      prior = c(prior(normal(0, 1), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 5,
      file = "G:\\My Drive\\Philippine rodents\\Chrotomyini_brms\\fits\\ch.36")
print(ch.36)
mcmc_plot(ch.36, "^b_", regex = T)
```

plot it by habitat:

```{r}
p.36 <- 
  ggplot(d, aes(x=mass_s, y = bvtv_s, color = hab_simp)) + 
  geom_point(pch = 16, cex = 4) +
  scale_color_manual(values = habmorecol)
p.36

```


Some models considering variation in relation to habitat!

##############################
#### Bone volume fraction ####
##############################

```{r}
ch.31 <- 
  brm(data = d, 
      family = gaussian,
      bvtv_s ~ 0 + hab_simp,
      prior = c(prior(normal(0, 1), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 5,
      file = "G:\\My Drive\\Philippine rodents\\Chrotomyini_brms\\fits\\ch.31")
print(ch.31)
mcmc_plot(ch.31, "^b_", regex = T)
```

Honestly I think that including 5 specimens from montane-mossy forest and 2 from lowland-montane is probably not doing the model any favors. So I'm going to try excluding those. use dd (as made above).

Then run the model again and see how it looks:

```{r}
ch.32 <- 
  brm(data = dd, 
      family = gaussian,
      bvtv_s ~ 0 + hab_simp,
      prior = c(prior(normal(0, 1), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 5,
      file = "G:\\My Drive\\Philippine rodents\\Chrotomyini_brms\\fits\\ch.32")
print(ch.32)
mcmc_plot(ch.32, "^b_", regex = T)
```

VERY interesting. Lower bvtv in mossy, medium in montane, high in lowland. I expect this is partially driven by body size because the lowland stuff is very chrotomys-heavy.
```{r}
p.32 <- ch.32 %>%
  gather_draws(b_hab_simplowland,b_hab_simpmontane, b_hab_simpmossy) %>%
  ggplot(aes(y = .variable, x = .value)) +
  stat_halfeye(aes(fill = .variable), 
               point_fill = "#000000", 
               shape = 21, 
               point_size = 3, 
               point_color = "#FFFFFF",
               interval_size = 7,
               interval_color = "grey40",
               .width = .89) +
  scale_fill_manual(values = habcol)+
  theme(legend.position = "none")+
  geom_vline(xintercept = 0, linetype = "dashed")+
  ggtitle(label = "Bone volume fraction by habitat only")
```


```{r}
ch.33 <- 
  brm(data = dd, 
      family = gaussian,
      bvtv_s ~ 0 + hab_simp + mass_s,
      prior = c(prior(normal(0, 1), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 5,
      file = "G:\\My Drive\\Philippine rodents\\Chrotomyini_brms\\fits\\ch.33")
print(ch.33)
mcmc_plot(ch.33, "^b_", regex = T)
```

The range of uncertainty in the lowland sample is pretty high - probably because it's only Apomys and Chrotomys mindorensis. Also it has the fewest specimens. However - the mossy and montane signals look very different EVEN when body size is accounted for!!

A little halfeye plot so we can see a more explicit visualization of the location of probability density, especially for that lowland group:

```{r}
p.33 <- ch.33 %>%
  gather_draws(b_hab_simplowland,b_hab_simpmontane, b_hab_simpmossy) %>%
  ggplot(aes(y = .variable, x = .value)) +
  stat_halfeye(aes(fill = .variable), 
               point_fill = "#000000", 
               shape = 21, 
               point_size = 3, 
               point_color = "#FFFFFF",
               interval_size = 7,
               interval_color = "grey40",
               .width = .89) +
  scale_fill_manual(values = habcol)+
  theme(legend.position = "none")+
  geom_vline(xintercept = 0, linetype = "dashed")+
  ggtitle(label = "Bone volume fraction by habitat, incl. mass")
```

Some checks for overly influential points:

```{r}
ch.33 <- add_criterion(ch.33, c("loo", "waic"))
loo(ch.33) # only one k above 0.5 - much better than I expected!
tibble(k   = ch.33$criteria$loo$diagnostics$pareto_k,
       row = 1:60,
       specimen = paste(d$specno[row], d$taxon[row])) %>% 
  arrange(desc(k))
```

The highest pareto k is Apomys sierrae. In fact, the top 5 specimens when listed in order of k are lowland specimens, and all of the top ten are either Apomys sierrae or C. whiteheadi/mindorensis.

Try with a Student distribution instead (less surprised by outliers) and see if it's overfit.

```{r}
ch.34 <- 
  brm(data = dd, 
      family = student,
      bvtv_s ~ 0 + hab_simp + mass_s,
      prior = c(prior(normal(0, 1), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      file = "G:\\My Drive\\Philippine rodents\\Chrotomyini_brms\\fits\\ch.34")
print(ch.34)
mcmc_plot(ch.34, "^b_", regex = T)

ch.34 <- add_criterion(ch.34, c("loo", "waic"))
loo_compare(ch.33, ch.34, criterion = "waic")%>% 
  print(simplify = F)
```

They are essentially the same, it looks like. Let's stick to gaussian dists unless we have any issues. 

Add phylo:

```{r}
ch.35 <- 
  brm(data = dd, 
      family = gaussian,
      bvtv_s ~ 0 + hab_simp + mass_s + (1|gr(phylo, cov = ch)) + (1|taxon),
      control = list(adapt_delta = 0.92), #inserted to decrease the number of divergent transitions here
      prior = c(prior(normal(0, 1), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      data2 = list(ch = ch),
      file = "G:\\My Drive\\Philippine rodents\\Chrotomyini_brms\\fits\\ch.35")
print(ch.35)
mcmc_plot(ch.35, "^b_", regex = T)
plot(ch.35, N = 3, ask = F)
```

```{r}
p.35 <- ch.35 %>%
  gather_draws(b_hab_simplowland,b_hab_simpmontane, b_hab_simpmossy) %>%
  ggplot(aes(y = .variable, x = .value)) +
  stat_halfeye(aes(fill = .variable), 
               point_fill = "#000000", 
               shape = 21, 
               point_size = 3, 
               point_color = "#FFFFFF",
               interval_size = 7,
               interval_color = "grey40",
               .width = .89) +
  scale_fill_manual(values = habcol)+
  theme(legend.position = "none")+
  geom_vline(xintercept = 0, linetype = "dashed")+
  ggtitle(label = "Bone volume fraction by habitat with mass and phylo")
```


```{r}
comp.35 <- ch.35 %>%
  gather_draws(b_hab_simplowland,b_hab_simpmontane, b_hab_simpmossy) %>%
  compare_levels(.value, by = .variable) %>%
  ungroup() %>%
  mutate(hab_simp = reorder(.variable, .value)) %>%
  ggplot(aes(y = .variable, x = .value)) +
  stat_halfeye() +
  geom_vline(xintercept = 0, linetype = "dashed")
```

Let's make a set of plots to compare.

```{r}
p.36/p.33|p.32/p.35
```

##############################
#### Trabecular thickness ####
##############################

By habitat only:
```{r}
ch.37 <- 
  brm(data = dd, 
      family = gaussian,
      tbth_s ~ 0 + hab_simp,
      prior = c(prior(normal(0, 1), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      file = "G:\\My Drive\\Philippine rodents\\Chrotomyini_brms\\fits\\ch.37")
print(ch.37)
mcmc_plot(ch.37, "^b_", regex = T)
```

By habitat and accounting for mass:

```{r}
ch.38 <- 
  brm(data = dd, 
      family = gaussian,
      tbth_s ~ 0 + hab_simp+mass_s,
      prior = c(prior(normal(0, 1), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 5,
      file = "G:\\My Drive\\Philippine rodents\\Chrotomyini_brms\\fits\\ch.38")
print(ch.38)
mcmc_plot(ch.38, "^b_", regex = T)
```

By habitat, accounting for mass + addition of phylogenetic covariation/intraspec:
```{r}
ch.39 <- 
  brm(data = dd, 
      family = gaussian,
      tbth_s ~ 0 + hab_simp + mass_s + (1|gr(phylo, cov = ch)) + (1|taxon),
      control = list(adapt_delta = 0.92), #inserted to decrease the number of divergent transitions here
      prior = c(prior(normal(0, 1), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      data2 = list(ch = ch),
      file = "G:\\My Drive\\Philippine rodents\\Chrotomyini_brms\\fits\\ch.39")
print(ch.39)
mcmc_plot(ch.39, "^b_", regex = T)
```

Plots:

```{r}
p.37 <- ch.37 %>%
  gather_draws(b_hab_simplowland,b_hab_simpmontane, b_hab_simpmossy) %>%
  ggplot(aes(y = .variable, x = .value)) +
  stat_halfeye(aes(fill = .variable), 
               point_fill = "#000000", 
               shape = 21, 
               point_size = 3, 
               point_color = "#FFFFFF",
               interval_size = 7,
               interval_color = "grey40",
               .width = .89) +
  scale_fill_manual(values = habcol)+
  theme(legend.position = "none")+
  geom_vline(xintercept = 0, linetype = "dashed")+
  ggtitle(label = "Tb thickness by habitat only")

p.38 <- ch.38 %>%
  gather_draws(b_hab_simplowland,b_hab_simpmontane, b_hab_simpmossy) %>%
  ggplot(aes(y = .variable, x = .value)) +
  stat_halfeye(aes(fill = .variable), 
               point_fill = "#000000", 
               shape = 21, 
               point_size = 3, 
               point_color = "#FFFFFF",
               interval_size = 7,
               interval_color = "grey40",
               .width = .89) +
  scale_fill_manual(values = habcol)+
  theme(legend.position = "none")+
  geom_vline(xintercept = 0, linetype = "dashed")+
  ggtitle(label = "Tb thickness by habitat + mass")

p.39 <- ch.39 %>%
  gather_draws(b_hab_simplowland,b_hab_simpmontane, b_hab_simpmossy) %>%
  ggplot(aes(y = .variable, x = .value)) +
  stat_halfeye(aes(fill = .variable), 
               point_fill = "#000000", 
               shape = 21, 
               point_size = 3, 
               point_color = "#FFFFFF",
               interval_size = 7,
               interval_color = "grey40",
               .width = .89) +
  scale_fill_manual(values = habcol)+
  theme(legend.position = "none")+
  geom_vline(xintercept = 0, linetype = "dashed")+
  ggtitle(label = "Tb thickness by habitat + mass + phylo")

p.40 <- 
  ggplot(d, aes(x=mass_s, y = tbth_s, color = hab_simp)) + 
  geom_point(pch = 16, cex = 4) +
  scale_color_manual(values = habmorecol)

p.40/p.38|p.37/p.39

```

############################
#### Trabecular spacing ####
############################

By habitat only:
```{r}
ch.41 <- 
  brm(data = dd, 
      family = gaussian,
      tbsp_s ~ 0 + hab_simp,
      prior = c(prior(normal(0, 1), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      file = "G:\\My Drive\\Philippine rodents\\Chrotomyini_brms\\fits\\ch.41")
print(ch.41)
mcmc_plot(ch.41, "^b_", regex = T)
```

By habitat and accounting for mass:

```{r}
ch.42 <- 
  brm(data = dd, 
      family = gaussian,
      tbsp_s ~ 0 + hab_simp+mass_s,
      prior = c(prior(normal(0, 1), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 5,
      file = "G:\\My Drive\\Philippine rodents\\Chrotomyini_brms\\fits\\ch.42")
print(ch.42)
mcmc_plot(ch.42, "^b_", regex = T)
```

By habitat, accounting for mass + addition of phylogenetic covariation/intraspec:
```{r}
ch.43 <- 
  brm(data = dd, 
      family = gaussian,
      tbsp_s ~ 0 + hab_simp + mass_s + (1|gr(phylo, cov = ch)) + (1|taxon),
      control = list(adapt_delta = 0.95), #inserted to decrease the number of divergent transitions here
      prior = c(prior(normal(0, 1), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      data2 = list(ch = ch),
      file = "G:\\My Drive\\Philippine rodents\\Chrotomyini_brms\\fits\\ch.43")
print(ch.43)
mcmc_plot(ch.43, "^b_", regex = T)
```

Plots:

```{r}
p.41 <- ch.41 %>%
  gather_draws(b_hab_simplowland,b_hab_simpmontane, b_hab_simpmossy) %>%
  ggplot(aes(y = .variable, x = .value)) +
  stat_halfeye(aes(fill = .variable), 
               point_fill = "#000000", 
               shape = 21, 
               point_size = 3, 
               point_color = "#FFFFFF",
               interval_size = 7,
               interval_color = "grey40",
               .width = .89) +
  scale_fill_manual(values = habcol)+
  theme(legend.position = "none")+
  geom_vline(xintercept = 0, linetype = "dashed")+
  ggtitle(label = "Tb spacing by habitat only")

p.42 <- ch.42 %>%
  gather_draws(b_hab_simplowland,b_hab_simpmontane, b_hab_simpmossy) %>%
  ggplot(aes(y = .variable, x = .value)) +
  stat_halfeye(aes(fill = .variable), 
               point_fill = "#000000", 
               shape = 21, 
               point_size = 3, 
               point_color = "#FFFFFF",
               interval_size = 7,
               interval_color = "grey40",
               .width = .89) +
  scale_fill_manual(values = habcol)+
  theme(legend.position = "none")+
  geom_vline(xintercept = 0, linetype = "dashed")+
  ggtitle(label = "Tb spacing by habitat + mass")

p.43 <- ch.43 %>%
  gather_draws(b_hab_simplowland,b_hab_simpmontane, b_hab_simpmossy) %>%
  ggplot(aes(y = .variable, x = .value)) +
  stat_halfeye(aes(fill = .variable), 
               point_fill = "#000000", 
               shape = 21, 
               point_size = 3, 
               point_color = "#FFFFFF",
               interval_size = 7,
               interval_color = "grey40",
               .width = .89) +
  scale_fill_manual(values = habcol)+
  theme(legend.position = "none")+
  geom_vline(xintercept = 0, linetype = "dashed")+
  ggtitle(label = "Tb spacing by habitat + mass + phylo")

p.44 <- 
  ggplot(d, aes(x=mass_s, y = tbsp_s, color = hab_simp)) + 
  geom_point(pch = 16, cex = 4) +
  scale_color_manual(values = habmorecol)

p.44/p.42|p.41/p.43

```
#################################

Can we plot the mass term from all four of the models?

```{r}
m.bvtv <- ch.35 %>% 
  gather_draws(b_mass_s) %>% 
  mutate(metric = "bvtv")

m.tbth <- ch.39 %>% 
  gather_draws(b_mass_s) %>% 
  mutate(metric = "tbth")

m.tbsp <- ch.43 %>% 
  gather_draws(b_mass_s) %>% 
  mutate(metric = "tbsp")

mass_b_all <- bind_rows(m.bvtv, m.tbth, m.tbsp)

mass.pl <-   mass_b_all %>% 
  ggplot(aes(y = metric, x = .value)) +
  stat_halfeye(aes(fill = .variable), 
               point_fill = "#000000", 
               shape = 21, 
               point_size = 3, 
               point_color = "#FFFFFF",
               interval_size = 7,
               interval_color = "grey40",
               .width = .89) +
  scale_fill_manual(values = "#CB2F7F")+
  theme(legend.position = "none")+
  geom_vline(xintercept = 0, linetype = "dashed")+
  ggtitle(label = "estimated effect of mass")
mass.pl
```

Ok Let's do the same thing for the phylo_sd term for comparison.. 

```{r}
ph.bvtv <- ch.35 %>% 
  gather_draws(sd_phylo__Intercept) %>% 
  mutate(metric = "bvtv")

ph.tbth <- ch.39 %>% 
  gather_draws(sd_phylo__Intercept) %>% 
  mutate(metric = "tbth")

ph.tbsp <- ch.43 %>% 
  gather_draws(sd_phylo__Intercept) %>% 
  mutate(metric = "tbsp")

phylo_sd_all <- bind_rows(ph.bvtv, ph.tbth, ph.tbsp)

ph.pl <-   phylo_sd_all %>% 
  ggplot(aes(y = metric, x = .value)) +
  stat_halfeye(aes(fill = .variable), 
               point_fill = "#000000", 
               shape = 21, 
               point_size = 3, 
               point_color = "#FFFFFF",
               interval_size = 7,
               interval_color = "grey40",
               .width = .89) +
  scale_fill_manual(values = habcol)+
  theme(legend.position = "none")+
  geom_vline(xintercept = 0, linetype = "dashed")+
  ggtitle(label = "estimated phylo sd (NOT lambda!)")
ph.pl
```


Do one for phylogenetic signal:
```{r}
hyp <- "sd_phylo__Intercept^2 / (sd_phylo__Intercept^2 + sd_taxon__Intercept^2 + sigma^2) = 0"
hh.bvtv <- hypothesis(ch.35, hyp, class = NULL)
hh.tbth <- hypothesis(ch.39, hyp, class = NULL)
hh.tbsp <- hypothesis(ch.43, hyp, class = NULL)

hph.bvtv.pl <- ggplot() +
  stat_halfeye(aes(x = hh.bvtv$samples$H1), fill = "red", alpha = 0.5) +
  theme_bw() +
  xlim(0,1) +
  labs(y = "density", x = "lambda: bone volume fraction")

hph.tbth.pl <- ggplot() +
  stat_halfeye(aes(x = hh.tbth$samples$H1), fill = "orange", alpha = 0.5)+
  theme_bw() +
  xlim(0,1)+
  labs(y = "density", x = "lambda: trabecular thickness")

hph.tbsp.pl <- ggplot() +
  stat_halfeye(aes(x = hh.tbsp$samples$H1), fill = "turquoise", alpha = 0.5)+
  theme_bw() +
  xlim(0,1) +
  labs(y = "density", x = "lambda: trabecular spacing")

hph.bvtv.pl/hph.tbth.pl/hph.tbsp.pl
```

